---
title: "Cross-Study Analyses (Studies 1 - 5)"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: FALSE
---

# Cross-Study Introduction

This Markdown documents the process to analyze studies 1 - 3 (FKA Studies 3- 5). It was created on 3/10/2020 by BWY.

This HTML was last knitted on: `r Sys.time()`

# Set Up

## Packages and Libraries

You must run this section before you can run any other chunks.

```{r packages, echo = FALSE}
#Make sure all packages are installed

list.of.packages <- c("readr", "tidyr", "dplyr", "magrittr", "psych", "stringr", "effsize", "shiny", "readr", "lme4", "lmerTest", "kableExtra", "knitr", "ggplot2", "ggthemes", "cowplot", "pwr", "psych")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
```


```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

#turn off scientific notation
options(scipen=999)
```


```{r libraries}
#Load all libraries
library(readr)
library(tidyr)
library(dplyr)
library(magrittr)
library(effsize)
library(lme4) # for mixed effects models
library(lmerTest)
library(kableExtra)
library(knitr)#kable
library(pwr)

##Raincloud Plot Libraries
source("../R_rainclouds.R")
library(cowplot)
library(ggthemes)
library(ggplot2)

```


```{r raincloud plot}
#raincloud plot theme
raincloud_theme <- theme(
  text = element_text(size = 10),
  axis.title.x = element_text(size = 16),
  axis.title.y = element_text(size = 16, margin = margin(r = 20)),
  axis.text = element_text(size = 14),
  legend.title=element_blank(),
  legend.text=element_text(size=16),
  legend.position = "right",
  plot.title = element_text(lineheight=.8, face="bold", size = 16),
  panel.border = element_blank(),
  panel.grid.minor = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.major.y = element_line(colour = 'light gray', size=0.5, linetype='solid'),
  axis.line = element_blank(), 
  axis.ticks.y = element_blank())
```

## Data Import

NOTE: manually changed `participantid` to `participantID` in s4

```{r data import, include=FALSE}

s1_models_difference_df <- read_csv("data/s1_models_difference_df.csv") %>%
  mutate(study = "Study 1")

s2_models_difference_df <- read_csv("data/s2_models_difference_df.csv") %>%
  mutate(study = "Study 2")

s3_models_difference_df <- read_csv("data/s3_models_difference_df.csv") %>%
  mutate(study = "Study 3")

s4_models_difference_df <- read_csv("data/s4_models_difference_df.csv") %>%
  mutate(study = "Study 4")

s5_models_difference_df <- read_csv("data/s5_models_difference_df.csv")

```

```{r import notice_df, include=FALSE}

# merge
models_difference_df <- s1_models_difference_df %>%
  full_join(s2_models_difference_df) %>%
  full_join(s3_models_difference_df) %>%
  full_join(s4_models_difference_df) %>%
  full_join(s5_models_difference_df)

nrow(models_difference_df) # 860


```

# Graph literacy analysis

## Main Effect

Graph literacy somewhat the size of the truncation effect: higher graph literacy -> slightly larger (~ .1 point increase in truncation effect for 10 points graph literacy)

```{r graphliteracy model}

graphliteracymodel <- lmer(difference ~ graphliteracy_sum_rating + (1 | participantID) + ( 1 | study), models_difference_df)

summary(graphliteracymodel)

```

```{r}
graphliteracymodel2 <- lmer(difference ~ graphliteracy_sum_rating + subject_condition + (1 | participantID) + (1 | study), models_difference_df)

summary(graphliteracymodel2)

```


Truncation effect is mean rating for truncated graphs - mean rating for control graphs. 

# Figure 5

```{r}

ggplot(models_difference_df, aes(x = graphliteracy_sum_rating, y = difference))+
  geom_smooth(method = lm, color = "#510D73", fill = "#510D73") +
  geom_jitter(alpha = 0.4, fill = NA)+
  ylab("truncation effect")+
  xlab("graph literacy score")+
  facet_wrap(~study) + 
  raincloud_theme +
  theme(
    axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
    axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'),
    panel.background = element_rect(fill = "white"),
    strip.text = element_text(size = 12),
    strip.background =element_rect(fill="white") # facet wrap label background
    )

#ggsave('../figures/Figure5.png', plot = last_plot(), width = 6, height = 4)
#ggsave('../figures/Figure5.tiff', plot = last_plot(), width = 6, height = 4, device = "tiff", dpi=300, limitsize = FALSE)

```
