---
title: "Data Analysis for Study 2"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: FALSE
---


# Study 2 Introduction

This Markdown documents the process to analyze the data for Study 2 which looked at the impact of a warning at the begining of the experiment on the difference between the truncated and the control condition. 

Data was collected on December 11th, 2017 by using Amazon's Mechanical Turk for ditribution and Qualtrics as a survey platform. 

This HTML was last knitted on: `r Sys.time()`

## Set Up

### Packages and Libraries

You must run this section before you can run any other chunks.


```{r packages, echo = FALSE}
#Make sure all packages are installed
list.of.packages <- c("readr", "tidyr", "dplyr", "magrittr", "psych", "stringr", "effsize", "shiny", "readr", "lme4", "lmerTest", "kableExtra", "knitr", "ggplot2", "ggthemes", "cowplot", "pwr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
```


```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

#turn off scientific notation
options(scipen=999)
```

```{r libraries}
#Load all libraries
library(readr)
library(tidyr)
library(dplyr)
library(magrittr)
library(effsize)
library(lme4) # for mixed effects models
library(lmerTest)
library(kableExtra)
library(knitr)#kable
library(pwr)

##Raincloud Plot Libraries
source("../R_rainclouds.R")
library(cowplot)
library(ggthemes)
library(ggplot2)

```


```{r raincloud plot}
#raincloud plot theme
raincloud_theme <- theme(
  text = element_text(size = 10),
  axis.title.x = element_text(size = 16),
  axis.title.y = element_text(size = 16, margin = margin(r = 20)),
  axis.text = element_text(size = 14),
  legend.title=element_blank(),
  legend.text=element_text(size=16),
  legend.position = "right",
  plot.title = element_text(lineheight=.8, face="bold", size = 16),
  panel.border = element_blank(),
  panel.grid.minor = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.major.y = element_line(colour = 'light gray', size=0.5, linetype='solid'),
  axis.line = element_blank(), 
  axis.ticks.y = element_blank())
```


### Data Import

You must run this section before you can run any other chunks.

```{r data import}
rating_df <- read_csv("data/clean/s2_rating_df.csv")
demographic_df <- read_csv("data/clean/s2_demographic_df.csv")
graphliteracy_df <- read_csv("data/clean/s2_graphliteracy_df.csv")
debriefing_df <- read_csv("data/clean/s2_debriefing_df.csv")
timing_df <- read_csv("data/clean/s2_timing_df.csv")
```


### Exclusions

You can exclude subjects who did not get the correct answer in the exercise by changing Exclude_Exercise_Check to TRUE. The next time you run all the code, these participants will be excluded. 

```{r}
Exclude_Exercise_Check <- FALSE

participants_excluded <- rating_df %>% 
  filter(check == "wrong") %>% 
  group_by(participantID, check) %>% 
  summarise(n = n()) %>% 
  group_by(check) %>% 
  summarise(n = n())

if(Exclude_Exercise_Check == TRUE){rating_df <- rating_df %>% 
  filter(check == "right")
} else{rating_df <- rating_df}

if(Exclude_Exercise_Check == TRUE){demographic_df <- demographic_df %>% 
  filter(check == "right")
} else{demographic_df <- demographic_df}

```

For this report, **`r if(Exclude_Exercise_Check == FALSE){paste("no")} else{paste(participants_excluded$n)}`** participants are being excluded from analysis. 


```{r addtional data frames, eval = TRUE}
#Create data frames that will be used throughout

#Calculate subject means by condition 
##(mean rating for truncated and mean rating for non-truncated graphs)
subject_mean_df <- rating_df %>% 
  group_by(condition, participantID) %>% 
  summarise(subject_mean_rating = mean(rating)) %>% 
  arrange(participantID)

#Calculate subject difference rating
##(non-truncated mean rating - truncated mean rating)
subject_difference_df <- subject_mean_df %>% 
  spread(condition, subject_mean_rating) %>% 
  mutate(difference =  truncated - control)


#Calculate subject overall graph literacy score
##(sum of graph literacy items)
subject_graphliteracy_df <- graphliteracy_df %>%
  group_by(participantID) %>%
  summarise(graphliteracy_sum_rating = sum(rating)) %>% 
  mutate(subject_condition = 1)


#Create mixed effects df
##(combination of ratings, overall graph literacy scores and some demographic questions --- education, gender and age)
models_df <- full_join(full_join(subject_graphliteracy_df, demographic_df[, c("participantID", "dem_ed", "dem_gender", "dem_age")]), rating_df)


##Create mixed effects difference df
##(df with subject truncation effect scores (difference between truncated and control) + all demographics and graph literacy scores) 
models_difference_df <- left_join(left_join(subject_difference_df,subject_graphliteracy_df), demographic_df[, c("participantID", "dem_ed", "dem_gender", "dem_age")])
models_difference_df$dem_ed <- as.factor(models_difference_df$dem_ed)

##############
# @camila -- When we run the models with difference as a DV, we are actually not using mixed effects models (because there is only one row per participant for the difference and for graph literacy), so the names of the dataframes are misleading
##############

#Create trimmed timing df
##(trims anything past 2 standard deviations from each participant's individual mean for each condition)
trimmed_timing_df <- timing_df %>% 
  group_by(participantID, condition) %>% 
  mutate(avg = mean(time), stdev = sd(time)) %>%
  filter(time <= 2*stdev+avg) %>%
  as.data.frame()

#Create graph literacy and subject difference scores df
difference_graphliteracy_df <- left_join(subject_difference_df,subject_graphliteracy_df)
```



--------------

## Methods

### Participants


```{r general participant information}
n <- demographic_df %>% 
  summarise(n())
```


```{r general participant information gender}
gender <- demographic_df %>% 
  group_by(dem_gender) %>% 
  summarise(n = n())

gender
  
```


```{r general participant information age}
age <- demographic_df %>% 
  summarise(mean_age = mean(dem_age), sd_age = sd(dem_age), 
            median_age = median(dem_age), min_age= min(dem_age), 
            max_age = max(dem_age), range_age = max(dem_age)-min(dem_age))

age
```

`r text_spec(n, bold = T)` Mturk workers (`r text_spec(gender[gender$dem_gender == "female",]$n, bold = T)` women, `r text_spec(gender[gender$dem_gender == "other",]$n, bold = T)` non-binary; M~age~ = `r text_spec(round(age$mean_age, 2), bold = T)` years, SD~age~ = `r text_spec(round(age$sd_age, 2), bold = T)`) from the United States and whose previous task approval rate was equal to or exceeded 85% participated in this experiment.

```{r general participant information age education}
education <- demographic_df %>% 
  group_by(dem_ed) %>% 
  summarise (n = n()) %>%
  mutate(percentage = n / sum(n)*100)

education
```

`r text_spec(round(sum(education[education$dem_ed >= 5,]$percentage)), bold = T)`% of participants reported having at least a Bachelor’s degree. The sample also included a range of self-reported graph literacy.

```{r summary statistics for graph literacy}
#Cronbach's Alpha
library(psych)

alpha <- graphliteracy_df %>% 
  spread(graphliteracy_question, rating) %>% 
  select(-group) %>% 
  psych::alpha()

#Data Summary
subject_graphliteracy_summary <- subject_graphliteracy_df %>% 
  summarise(n = n(), 
            mean = mean(graphliteracy_sum_rating), 
            sd = sd(graphliteracy_sum_rating), 
            median = median(graphliteracy_sum_rating),
            min= min(graphliteracy_sum_rating), 
            max = max(graphliteracy_sum_rating), 
            range=max(graphliteracy_sum_rating)-min(graphliteracy_sum_rating)) %>% 
  mutate(alpha = round(alpha$total$std.alpha,2))

subject_graphliteracy_summary
```

 
### Power

```{r power analysis}
power <- pwr.t.test(d = 0.35, sig.level = 0.05, power = 0.90,  type = c("paired"))
```

Because we hypothesized that the truncation effect would be reduced with a warning, we determined our sample size a priori based on a power analysis predicting a small-to-medium effect size of `r text_spec(round(power$d,2), bold = T)`, an alpha of `r text_spec(power$sig.level, bold = T)`, and power of `r text_spec(power$power, bold = T)`, which estimated a required sample size of at least `r text_spec(round(power$n), bold = T)` participants. 


### Exercise Check

```{r}
exercise_check <- rating_df %>% 
  group_by(participantID, check) %>% 
  summarise(n()) %>% 
  group_by(check) %>% 
  summarise(n = n()) %>% 
  mutate(percentage = n / sum(n)*100)
```

`r text_spec(exercise_check[exercise_check$check == "wrong",]$n, bold = T)` did not get the manipulation check question right. Accuracy was `r text_spec(round(exercise_check[exercise_check$check == "right",]$percentage,2), bold = T)`%

For this report, **`r if(Exclude_Exercise_Check == FALSE){paste("no")} else{paste(participants_excluded$n)}`** participants are being excluded from analysis. If you want to see results when participants who got the exercise wrong are excluded/included, you can go to the section called Exclusions (at the top of this file) and change Exclude_Exercise_Check <- TRUE

Excluding the participants who did not answer the training exercise correctly initially does not change the pattern or results or conclusions drawn.

## Results

### Truncation effect

```{r summary statistics for graph ratings}
#Summarize Accross Subjects
subject_mean_summary <- subject_mean_df %>% 
  group_by(condition) %>% 
  summarise(n = n(), mean = mean(subject_mean_rating), 
            sd = sd(subject_mean_rating), 
            median = median(subject_mean_rating),
            min= min(subject_mean_rating), max = max(subject_mean_rating), 
            range =max(subject_mean_rating)-min(subject_mean_rating))

subject_mean_summary
  
```


Average ratings for truncated graphs (M~truncated~ = `r text_spec(round(subject_mean_summary[subject_mean_summary$condition == "truncated",]$mean,2), bold = T)`, SD~truncated~ = `r text_spec(round(subject_mean_summary[subject_mean_summary$condition == "truncated",]$sd,2), bold = T)`)  were higher than average ratings for control graphs (M~control~ = `r text_spec(round(subject_mean_summary[subject_mean_summary$condition == "control",]$mean,2), bold = T)`, SD~control~ = `r text_spec(round(subject_mean_summary[subject_mean_summary$condition == "control",]$sd,2), bold = T)`), as shown in Figure 3B. 

```{r figure 5 prep}

#Flat Violin Set Up

subject_mean_df %<>% 
  mutate(subject_condition = 1)

errbar_lims <- subject_mean_df %>% 
  group_by(condition, subject_condition) %>% 
  summarise(mean=mean(subject_mean_rating), se=sd(subject_mean_rating)/sqrt(n()), 
                        upper=mean+(2*se), lower=mean-(2*se))

```


```{r figure 5}
#Flat Violin Pink and Blue

ggplot(subject_mean_df, aes(x = subject_condition, y = subject_mean_rating, fill = condition)) +
  geom_flat_violin(aes(fill = condition),position = position_nudge(x = .1, y = 0), 
                   adjust = 1.5, trim = FALSE, colour = NA, alpha = 0.7)+
  geom_point(aes(x = as.numeric(subject_condition), y = subject_mean_rating, colour = condition),
             position = position_jitter(0.05), shape = 20, size = 2, alpha = 0.7)+
  geom_point(data=errbar_lims, aes(x=subject_condition, y=mean, color = condition, fill = condition), position = position_nudge(x = -0.15, y = 0))+
  geom_errorbar(data=errbar_lims, aes(x=subject_condition, y=mean, ymax=upper, ymin=lower, color = condition),stat='identity', size = 1, width=.06, position = position_nudge(x = -0.15, y = 0))+
  #Color
  scale_colour_manual(values = c("#000A77", "#FF6171"))+
  scale_fill_manual(values = c("#000A77", "#FF6171"))+
  #Axis
  coord_cartesian(ylim=c(1, 7), xlim =c(0.75,1.7)) + 
  scale_y_continuous(breaks=seq(1, 7, 1))+
  ylab("mean rating")+
  xlab("")+
  #Theme
  raincloud_theme+
  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())+
  guides(fill = guide_legend(reverse=TRUE), color = guide_legend(reverse=TRUE))+
  ggtitle("Figure 3B")

ggsave('../figures/figure3_B.png', plot = last_plot(), width = 6, height = 5)
ggsave('../figures/figure3_B.tiff', plot = last_plot(), width = 6, height = 5, device = "tiff")
```


```{r mean difference}
M_difference <- subject_difference_df %>% 
  summarise(mean = mean(difference))
```

```{r t test for graph ratings}
t_test <- t.test(subject_mean_rating ~ condition, subject_mean_df, paired = TRUE)

t_test

```

```{r effect size for graph ratings}
cohen_d <- effsize::cohen.d(subject_mean_df$subject_mean_rating, subject_mean_df$condition, paired = TRUE, na.rm = TRUE)

cohen_d
```


A paired t-test revealed that this difference was statistically significant with a `r text_spec(as.character(cohen_d$magnitude), bold = T)` effect size, t(`r text_spec(round(t_test$parameter,2), bold = T)`) = `r text_spec(abs(round(t_test$statistic,2)), bold = T)`, p < `r text_spec(format(round(t_test$p.value,3), nsmall= 2), bold = T)`, M~difference~ = `r text_spec(round(M_difference$mean,2), bold = T)`, d = `r text_spec(abs(round(cohen_d$estimate,2)), bold = T)`. These results suggest that despite the initial explanation and warning given, participants rated the differences between quantities expressed as bars in the truncated condition as being greater than the differences between quantities expressed as bars in the control condition.


```{r truncation effect direction}
truncation_effect_direction <- subject_difference_df %>% 
  mutate(direction = ifelse(difference > 0, "expected", "unexpected")) %>%
  group_by(direction) %>% 
  summarise(n= n()) %>% 
  mutate(percentage = n / sum(n)*100)

truncation_effect_direction           
```

`r text_spec(round(truncation_effect_direction[truncation_effect_direction$direction == "expected",]$percentage), bold = T)`% of participants showed a truncation effect in the expected direction i.e., for `r text_spec(round(truncation_effect_direction[truncation_effect_direction$direction == "expected",]$n), bold = T)` of our `r text_spec(n, bold = T)` participants, their ratings of comparisons shown by truncated graphs were larger than their ratings of comparisons shown by graphs without truncated vertical axes.



## Supplemental Information

### Methods (Supplemental Information)

#### Participants (Supplemental Information)

```{r supplemental participant information language}
langauge <- demographic_df %>% 
  group_by(dem_language) %>% 
  summarise(n = n())

other_language <- demographic_df %>% 
  group_by(dem_language_2_TEXT) %>% 
  summarise(n = n())
```


`r text_spec(langauge[1,]$n, bold = T)` participants reported English as their first language, `r text_spec(other_language[2,]$n, bold = T)` reported `r text_spec(other_language[2,]$dem_language_2_TEXT, bold = T)` and `r text_spec(other_language[3,]$n, bold = T)` reported `r text_spec(other_language[3,]$dem_language_2_TEXT, bold = T)` as their first language. 


```{r supplemental participant information education}
education <- demographic_df %>% 
  group_by(dem_ed) %>% 
  summarise (n = n()) %>%
  mutate(percentage = n / sum(n)*100)

education
   
```
`r text_spec(round(sum(education[education$dem_ed >= 5,]$percentage)), bold = T)`% of participants reported having at least a Bachelor’s degree. 



```{r supplemental participant information duration}
duration <- demographic_df %>% 
  summarise(mean_duration = mean(duration_min), 
            sd_duration = sd(duration_min), 
            median_duration = median(duration_min), 
            min_duration= min(duration_min), 
            max_duration = max(duration_min), 
            range_duration = max(duration_min)-min(duration_min))

duration
  
```


The experiment took participants an average of M~duration~ = `r text_spec(round(duration$mean_duration, 2), bold = T)` minutes, (SD~duration~ = `r text_spec(round(duration$sd_duration, 2), bold = T)` minutes). 



### Results (Supplemental Information)

#### Graph Literacy (Supplemental Information)


```{r graphliteracy model}
graphliteracymodel <- lm(difference ~ graphliteracy_sum_rating, models_difference_df)
summary_graphliteracymodel <- summary(graphliteracymodel)
summary_graphliteracymodel
```

As in Study 1, we explore whether participants’ graph literacy scores predict the how differently they rated truncated versus control graphs. As before, we did not find this model to be predictive: F(1, 107) = `r text_spec(round(summary_graphliteracymodel$fstatistic[1],4), bold = T)`, adjusted R2 = `r text_spec(round(summary_graphliteracymodel$adj.r.squared,3), bold = T)`, p = `r text_spec(round(summary_graphliteracymodel$coefficient[2,4],2), bold = T)`


Truncation effect is mean rating for truncated graphs - mean rating for control graphs. 
```{r}
ggplot(models_difference_df, aes(x = graphliteracy_sum_rating, y = difference))+
  geom_smooth(method = lm, color = "#510D73", fill = "#510D73") +
  geom_jitter(alpha = 0.7)+
  ylab("truncation effect")+
  xlab("graph literacy score")+
  ggtitle("Figure SI2 Graph Literacy")+
  raincloud_theme+
  theme(axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'))

ggsave('../figures/figureSI2_graphliteracy.png', plot = last_plot(), width = 6, height = 4)
```


#### Education (Supplemental Information)

Education Anova
```{r}
# education Only
educationOnly <- aov(difference ~ dem_ed, models_difference_df)
summary(educationOnly)
```

Education and graph literacy 
```{r}
models_difference_df$dem_ed <- as.numeric(models_difference_df$dem_ed)

education_graphliteracy <- lm(graphliteracy_sum_rating ~ dem_ed, models_difference_df)
summary(education_graphliteracy)
```


#### Age (Supplemental Information)

Age Linear Regression
```{r}
#age Only
ageOnly<- lm(difference ~ dem_age, models_difference_df)
summary(ageOnly)
```

Figure depicts this age null relationship.

```{r figure SI2 age}
ggplot(models_difference_df, aes(x = dem_age, y = difference))+
  geom_smooth(method = lm, color = "#510D73", fill = "#510D73") +
  geom_jitter(alpha = 0.7)+
  ylab("truncation effect")+
  xlab("age")+
  #ggtitle(Figure SI2 age")+
  raincloud_theme+
  theme(axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'))

ggsave('../figures/figureSI2_age.png', plot = last_plot(), width = 6, height = 4)

```

#### Item Analysis (Supplemental Information)

Here we summarise the data for each item. This is helpful to identify if the effect is driven by a few graphs or if it is seen across all graphs in the materials. 

```{r item analysis}
item_mean_df <- rating_df %>% 
  group_by(question,condition) %>% 
  summarise(item_mean_rating = mean(rating)) %>% 
  arrange(question)

#Data Summary
item_mean_summary <- item_mean_df %>% 
  group_by(condition) %>% 
  summarise(n = n(), mean = mean(item_mean_rating), sd = sd(item_mean_rating), 
            median = median(item_mean_rating), min= min(item_mean_rating), 
            max = max(item_mean_rating), 
            range = max(item_mean_rating)-min(item_mean_rating))

item_mean_summary
  
```

```{r}
item_mean_df %>% 
  spread(condition, item_mean_rating) %>%
  mutate(diff=truncated - control)
```

```{r}
item_mean_summary <- item_mean_df %>% 
  spread(condition, item_mean_rating) %>%
  mutate(diff=truncated - control) %>% 
  ungroup() %>% 
  summarise(n = n(), mean = mean(diff), sd = sd(diff), 
            median = median(diff), min= min(abs(diff)), 
            max = max(abs(diff)), 
            range = max(abs(diff))-min(abs(diff)))
```


```{r}
ggplot(rating_df, aes(x = condition, y = rating, group = condition, color = condition))+
  geom_jitter(width = 0.23, alpha = 0.5)+
  stat_summary(fun.y = mean, geom = "point", color = "black") + 
  stat_summary(fun.y=mean, color="black", geom="line", aes(group = 1))+
  facet_wrap(~as.factor(question))+
  scale_colour_brewer(palette = "Set1", direction=-1)

ggsave("../figures/figureSI2_item_analysis.png", device = "png", width = 14, height = 8, plot = last_plot(),scale = 1)
```

```{r}
item_mean_df %>% 
  spread(condition, item_mean_rating) %>%
  mutate(diff=truncated - control)%>% 
  filter(diff <0)
```

```{r}
item_mean_summary
```


The truncation effect was observed for all graphs except graph 14, 23. On average, the 7-point ratings were `r text_spec(round(item_mean_summary$mean,2), bold = T)` (SD = `r text_spec(round(item_mean_summary$sd,2), bold = T)`) higher when graphs were seen in the truncated condition than when they were seen in the control condition. The maximum average absolute difference between the trunccted and control conditions was for graph 33 M = `r text_spec(round(item_mean_summary$max,2), bold = T)` The minimum average absolute difference was for graph 23 where M = `r text_spec(round(item_mean_summary$min,2), bold = T)`.


#### Timing (Supplemental Information)

```{r supplemental information timing}

Trimming <- TRUE

if(Trimming == TRUE){timing_t_df = trimmed_timing_df
} else{timing_t_df = timing_df}
```


Timing information is in seconds. 

You can choose to trim or not timing data. By default any timing that is 2 standard deviations away from the mean (for each participant for each condition) is trimmed. 

For this report, any timing that is 2 standard deviations away from the mean (for each participant for each condition) `r if(Trimming == TRUE){paste("WAS")} else{paste("WAS NOT")}` trimmed.  If you want to see results when participants who got the exercise wrong are excluded/included, you can go to the section called Exclusions (at the top of this file) and change Trimming <- FALSE

```{r}
subject_timing <-  timing_t_df %>%
  group_by(participantID, condition) %>%
  summarise(subject_mean_timing = mean(time))

#Data Summary
subject_timing_summary <- subject_timing %>% 
  group_by(condition) %>% 
  summarise(n = n(), mean = mean(subject_mean_timing), 
            sd = sd(subject_mean_timing), 
            median = median(subject_mean_timing), min= min(subject_mean_timing), 
            max = max(subject_mean_timing), 
            range = max(subject_mean_timing)-min(subject_mean_timing))

subject_timing_summary
  
```



Item Analysis for Timing
```{r}
item_timing_df <- timing_t_df %>%
  group_by(question, condition) %>%
  summarise(item_mean_timing = mean(time))


#Data Summary
item_timing_summary <- item_timing_df %>% 
  group_by(condition) %>% 
  summarise(n = n(), mean = mean(item_mean_timing), sd = sd(item_mean_timing), 
            median = median(item_mean_timing), min= min(item_mean_timing), 
            max = max(item_mean_timing), 
            range = max(item_mean_timing)-min(item_mean_timing))

item_timing_summary
  
```


```{r}
ggplot(timing_t_df, aes(x = condition, y = time, group = condition, color = condition))+
  geom_jitter(width = 0.23, alpha = 0.5)+
  stat_summary(fun.y = mean, geom = "point", color = "black") + 
  stat_summary(fun.y=mean, color="black", geom="line", aes(group = 1))+
  facet_wrap(~as.factor(question))+
  scale_colour_brewer(palette = "Set1", direction=-1)
```





